Medical Concept Extraction

### 2012-12-17 paper/paper.tex
updated references
first stab at discussion

In the work, we present a system for extracting problems, tests, and treatments from medical records at comparable performance (approximately 0.8 F1 score on average in testing) to existing methods. Our classification strategies prove successful in spite our relatively limited data set, which is comprised of hospital records from two Boston area facilities. Our portable open-source system is implemented in python with easily configurable features and three classifier options.

@@

\subsubsection{Porter Stemmer}
We already discussed the benefits of using entire words as features, but similar words come in many different forms. Motivated by generalization, we chose to use word stems as features and approached this feature extraction using three different stemming algorithms--two at the word-level and one at the sentence level (See 5.3.2). The addition of word stem features to our system improved our testing CRF F1 scores from 0.58 to 0.82. We used built-in NLTK implementations of all of the stemming algorithms described in future sections.

@@

\subsubsection{Word Shape}
In the same spirit of our word length and MIT RegEx features, we decided to look at the high-level shapes of individual words for better generalization. More specifically, we encode words into new strings that give a sense of their capitalization and punctuation patterns as well as approximate length. We tested a variety of open-sourced word shape classification methods including two methods written by Dan Klein, one of which simply classifies words as numeric, strictly upper or lower cased, or mixed, and a second, more fine-grained classifier that attempts to group words into equivalence classes by collapsing sequences of the same type. Additional classifiers developed by Christopher Manning seek to make less distinctions mid-sequence by sorting words based on length. Each word shape classifier result added an additional dimension to our feature space. By naively including all word shape alternatives as features, we were able to improve our testbed CRF F1 score from 0.82 to 0.85.

@@

Talk about what we saw in terms of which was the best feature set and model.

Be sure to include examples of cool cases where it caught a difficult label

Be sure to include FAILURE cases

We stopped using libsvms.

>>>>

In the work, we present a system for extracting problems, tests, and treatments from medical records at comparable performance (approximately 0.8 F1 score on average in testing) to existing methods. Our classification strategies prove successful in spite our relatively limited data set, which is comprised of hospital records from two Boston area facilities. Our portable open-source system is implemented in python with easily configurable and extendable feature set and three classifier options.

@@

\subsubsection{Porter Stemmer}
We already discussed the benefits of using entire words as features, but similar words come in many different forms. Motivated by generalization, we chose to use word stems as features and approached this feature extraction using three different stemming algorithms--two at the word-level and one at the sentence level (See 5.3.2). The addition of word stem features to our system improved our toy test CRF F1 scores from 0.58 to 0.82. We used built-in NLTK implementations of all of the stemming algorithms described in future sections.

@@

\subsubsection{Word Shape}
In the same spirit of our word length and MIT RegEx features, we decided to look at the high-level shapes of individual words for better generalization. More specifically, we encode words into new strings that give a sense of their capitalization and punctuation patterns as well as approximate length. We tested a variety of open-sourced word shape classification methods including two methods written by Dan Klein, one of which simply classifies words as numeric, strictly upper or lower cased, or mixed, and a second, more fine-grained classifier that attempts to group words into equivalence classes by collapsing sequences of the same type. Additional classifiers developed by Christopher Manning seek to make less distinctions mid-sequence by sorting words based on length. Each word shape classifier result added an additional dimension to our feature space. By naively including all word shape alternatives as features, we were able to improve our toy testbed CRF F1 score from 0.82 to 0.85.

@@

As priorly mentioned, we tested our feature set using three classifiers in search of the best concept extraction system (general SVM, linear SVM, and CRF), all of which yielded comparable results in terms of their average F1 scores, but showed significant difference in terms of efficiency. Our less efficient classifier was eventually eliminated from our system, as discussed in the subsequent section.

After extensive testing of various subsets from our initial feature set, we decided to include all of the features discussed in Section 4.2 as the complete sent proved to be optimal for classification performance. Certain features (metric units, word endings, prognosis locations, and flag words as discussed in Section 4.2.7) were eliminated before significant testing was performed as their proved to have an appreciable negative impact on our F1 scores when testing toy examples.
